Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-01-17 Multiresolution_framework/ltx/manuscript.tex
working on MRA SS model section some more some more

The multi-resolution approximation \cite{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering an $n$-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(\mathbf{r}\right)=\sum_{\mathbf{l} \in \mathbb{Z}^n}\alpha_{j_0,\mathbf{l}} \underbrace{\phi_{j_0,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{scaling}} + \sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l} \in \mathbb{Z}^n}\beta_{j,\mathbf{l}} \underbrace{\psi_{j,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{wavelet}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,\mathbf{l}}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,\mathbf{l}}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,\mathbf{l}}\left(\mathbf{r}\right) =& 2^{\frac{j}{2}}\phi\left(2^j\mathbf{r}-\mathbf{l}\right) \\	
\psi_{j,\mathbf{l}}\left(\mathbf{r}\right) =& 2^{\frac{j}{2}}\psi\left(2^j\mathbf{r}-\mathbf{l}\right).
\end{align}
The parameters $j$ and $\mathbf{l}$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THEAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

@@

% where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, \dean{($\boldsymbol\theta \in \mathbb{R}^{n_{\theta}\times (J+2)}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x \times (J+2)}$)} are defined as 
% \begin{align}
% \boldsymbol\theta^\top \triangleq& [\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
% \label{KernelWeights} \\
% \mathbf{x}_{t}^\top \triangleq& [\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}].
% \label{FieldWeights}
% \end{align}
% The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, \mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n \right\rbrace $ and $\left\lbrace x_{t,j_0, \mathbf{l}}: \mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$, contain all the coefficients $\left\lbrace \beta_{j,\mathbf{l}} :\mathbf{l} \in \mathbb{Z}^n\right\rbrace$ and $\left\lbrace  \check x_{t,j,\mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively.
% 
% The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
% \begin{align}
%     \label{KernelBasisVector}
%     \boldsymbol\lambda^\top(\mathbf{r}) \triangleq& \left[
%     \begin{array}{ccccc} \boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
%     \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
%     \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
%     \cdots &
%     \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right] \\
%     \label{FieldBasisVector}
%     \boldsymbol\mu^\top (\mathbf{r}) \triangleq& \left[
%     \begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
%     \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
%     \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
%     \cdots &
%     \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right]
% \end{align}
% 
% The vectors in  (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in  (\ref{KernelWeights}) and (\ref{FieldWeights}).  
Now we have the compact for of the decomposition we can substitute this into the neural field equation \eqref{eq:DiscreteTimeModel} giving
\begin{equation}\label{eq:DiscreteTimeModel}
	\boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t = 
	\xi \boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t + 
	T_s \int_\Omega { 
	    \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r}'\right)
	    f\left(\boldsymbol\mu^\top\left(\mathbf{r}'\right)\mathbf{x}_t\right) 
	\, \mathrm{d}\mathbf{r}'}  
	+ e_t\left(\mathbf{r}\right), 
\end{equation}

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
\end{align}
The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}\label{eq:CovMatrix}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}
\section{B-spline MRAIDE in State-Space }
B-spline functions are one of the most basic elements for wavelet construction, these functions are composed of  piecewise polynomials combined smoothly at some breaking points (knots) which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \cite{Goswami1999} where the higher the order, the smoother the resulting B-spline becomes (see Figure~\ref{fig:Figure0}).
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure0.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (coloured lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

>>>>

The multi-resolution approximation \cite{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering an $n$-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(\mathbf{r}\right)=\sum_{\mathbf{l} \in \mathbb{Z}^n}\alpha_{j_0,\mathbf{l}} \underbrace{\phi_{j_0,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{scaling}} + \sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l} \in \mathbb{Z}^n}\beta_{j,\mathbf{l}} \underbrace{\psi_{j,\mathbf{l}}\left(\mathbf{r}\right)}_{\text{wavelet}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,\mathbf{l}}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,\mathbf{l}}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,\mathbf{l}}\left(\mathbf{r}\right) =& 2^{\frac{j}{2}}\phi\left(2^j\mathbf{r}-\mathbf{l}\right) \\	
\psi_{j,\mathbf{l}}\left(\mathbf{r}\right) =& 2^{\frac{j}{2}}\psi\left(2^j\mathbf{r}-\mathbf{l}\right).
\end{align}
The parameters $j$ and $\mathbf{l}$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

@@

% where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, \dean{($\boldsymbol\theta \in \mathbb{R}^{n_{\theta}\times (J+2)}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x \times (J+2)}$)} are defined as 
% \begin{align}
% \boldsymbol\theta^\top \triangleq& [\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
% \label{KernelWeights} \\
% \mathbf{x}_{t}^\top \triangleq& [\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}].
% \label{FieldWeights}
% \end{align}
% The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, \mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n \right\rbrace $ and $\left\lbrace x_{t,j_0, \mathbf{l}}: \mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$, contain all the coefficients $\left\lbrace \beta_{j,\mathbf{l}} :\mathbf{l} \in \mathbb{Z}^n\right\rbrace$ and $\left\lbrace  \check x_{t,j,\mathbf{l}}:\mathbf{l} \in \mathbb{Z}^n\right\rbrace$, respectively.
% 
% The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
% \begin{align}
%     \label{KernelBasisVector}
%     \boldsymbol\lambda^\top(\mathbf{r}) \triangleq& \left[
%     \begin{array}{ccccc} \boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
%     \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
%     \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
%     \cdots &
%     \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right] \\
%     \label{FieldBasisVector}
%     \boldsymbol\mu^\top (\mathbf{r}) \triangleq& \left[
%     \begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(\mathbf{r}) &
%     \boldsymbol\psi_{j_0}^\top(\mathbf{r}) & 
%     \boldsymbol\psi_{j_0+1}^\top(\mathbf{r}) &
%     \cdots &
%     \boldsymbol\psi_{J}^\top(\mathbf{r})\end{array}\right]
% \end{align}
% 
% The vectors in  (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in  (\ref{KernelWeights}) and (\ref{FieldWeights}).  
Now we have the compact for of the decomposition we can substitute this into the neural field equation \eqref{eq:DiscreteTimeModel} giving
\begin{equation}\label{eq:ApproxDiscreteTimeModel}
	\boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t = 
	\xi \boldsymbol\mu^\top\left(\mathbf{r}\right)\mathbf{x}_t + 
	T_s \int_\Omega { 
	    \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r}'\right)
	    f\left(\boldsymbol\mu^\top\left(\mathbf{r}'\right)\mathbf{x}_t\right) 
	\, \mathrm{d}\mathbf{r}'}  
	+ e_t\left(\mathbf{r}\right).
\end{equation}
The aim of the next part of the derivation is to isolate the state vector on the left hand side of the equation. To do this we first multiply both sides of equation~\eqref{eq:ApproxDiscreteTimeModel} by $\boldsymbol\mu\left(r\right)$ and integrate over space giving (\dean{UP TO HERE})
\begin{equation}\label{eq:ApproxDiscreteTimeModel2}
	\int_{\Omega} \boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(\mathbf{r}\right) \mathrm{d}\mathbf{r} \mathbf{x}_t = 
	\xi \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) \mathrm{d}\mathbf{r} \mathbf{x}_t + 
	T_s \int_{\Omega}\boldsymbol\mu\left(r\right)\int_\Omega { 
	    \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r}'\right)
	    f\left(\boldsymbol\mu^\top\left(\mathbf{r}'\right)\mathbf{x}_t\right) 
	\, \mathrm{d}\mathbf{r}'\mathrm{d}\mathbf{r}}  
	+ \int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(\mathbf{r}\right)\mathrm{d}\mathbf{r}.
\end{equation}

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) \mathrm{d}\mathbf{r} \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
\end{align}
The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}\label{eq:CovMatrix}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}
\section{B-spline MRAIDE in State-Space }
B-spline functions are one of the most basic elements for wavelet construction, these functions are composed of  piecewise polynomials combined smoothly at some breaking points (knots) which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \cite{Goswami1999} where the higher the order, the smoother the resulting B-spline becomes (see Figure~\ref{fig:Figure0}).
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure0.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (coloured lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}
